---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this.  -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


# pitchR  <img src= "https://github.com/Reed-Math241/pkgGrpq/blob/master/figs/IMG_0175.png" align="right" width=175 />


<!-- badges: start -->
<!-- badges: end -->


The goal of pitchR is to create an accessible dataset with advanced pitcher statistics by season and their corresponding salaries.  This dataset contains statistics from the 2018 thru 2020 seasons.

## Installation

The development version of XXX is available from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Reed-Math241/pkgDemo")
```

## About the Data

Data were collected and made available by [Spotrac](https://www.spotrac.com/mlb/payroll/) and the [Savant](https://baseballsavant.mlb.com/statcast_search).

The pitchR package contains one dataset, with data on Pitchers from the 2018-2020 Seasons. 

```{r showdata}
library(pitchR)
data(package = 'pitchR')
``` 

Here is a simplified version of the raw data; see `?pitchR` for more info: 

```{r example-pitchR}
head(pitchR)
```

## Example







