---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this.  -->

```{r, include = FALSE}

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(pitchR)
```


# pitchR  <img src= "https://github.com/Reed-Math241/pkgGrpq/blob/master/figs/IMG_0175.png" align="right" width=175 />


<!-- badges: start -->
<!-- badges: end -->


The goal of pitchR is to create an accessible dataset with advanced pitcher statistics by season and their corresponding salaries.  This dataset contains statistics from the 2018, 2019, and 2020 seasons.

## Installation

The development version of pitchR is available from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Reed-Math241/pkgDemo")
```

## About the Data

Data were collected and made available by [Spotrac](https://www.spotrac.com/mlb/payroll/) and the [Savant](https://baseballsavant.mlb.com/statcast_search).

The pitchR package contains one dataset, with data on Pitchers from the 2018-2020 Seasons. 

```{r}
library(pitchR)
data('pitchR')
``` 

Here is a simplified version of the raw data; see `?pitchR` for more info: 

```{r}
head(pitchR)
```



```{r}
ggplot(data = pitchR, aes(x = velocity, y = spin_rate)) +
  geom_point(
    aes(color = whiffs),
    alpha = 0.4
    ) +
  theme_minimal() +
  scale_color_distiller(direction = -1, palette = "YlGn") +
  labs(
    title = "Velocity, Spin Rate, and Whiffs",
    subtitle = "Whiffs: Total Number of times a batter swung and whiffed on the player's pitch",
    x = "Velocity (MPH)",
    y = "Spin Rate (Revolutions per Min.)"
    ) +
  theme(
    legend.background = element_rect(fill = "white", color = NA),
    plot.title.position = "plot",
    plot.caption = element_text(hjust = 0, face = "italic"),
    plot.caption.position = "plot"
    )
```









